Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-12 paper/type-inference.tex
Type Inference complete.

>>>>

\subsection{Higher-Order Type Inference}

We have so far ignored function types in the context of type
inference. The region type system for $\FB$ admits function types with
higher-rank region polymorphism, allowing region variables to be
quantified in the types of higher-order arguments. The type system
clearly does not have the principal typing property, as demonstrated
by the \C{app2} method below:
\begin{codejava}
bool app2(A x, A y, (A $\rightarrow$ bool) f) 
    { return (f x && f y); }
\end{codejava}
There are two possibilities for the region type of \C{app2}: one,
where it allows \C{x} and \C{y} to be allocated in different regions,
but requires \C{f} to be region-polymorphic (i.e, \C{f} should accept
arguments from any region), and second, where it requires \C{x} and
\C{y} to be allocated in the same region, and allows \C{f} to be
monomorphic. None of the two types are more general than the other,
and it is not clear which one is more preferable. The concrete
manifestation of this dilemma is a constraint of form 
$\isvalid{\phictxt \conj \varphi_1}{\phicstr \conj \varphi_2}$, with
no additional constriants to help determine $\varphi_1$ or $\varphi_2$
before solving this constraint. Since it is always possible to
strengthen $\varphi_2$ and then find a suitable $\varphi_1$, there
does not exist unique solution to such constraints. Nonetheless, it is
possible to bias the constraint-solving algorithm to consistently
prefer one solution over the other.

In context of \name, however, we adopt a trivial solution of
assigning the same region to the arguments and return values in
(higher-order) function types. For user-defined functions (e.g.,
predicate in the \C{WHERE} clause of the \C{SELECT} dataflow query)
and lambda arguments to the $\RgnZ$ class constructor, which are the
prominent uses of higher-order functions in \name, the simplistic
approach seems to suffice in practice. However, it is always possible
that this approach makes type checker reject a program that is
otherwise valid. In such cases, programmer can help the type checker
by writing region annotations for only the higher-order functions.

% \begin{center}
%   \C{app2} : $\inang{\rho_0,\rho_1}$(\C{A}$\inang{\rho_0}$, 
%               \C{A}$\inang{\rho_0}$,
%               $\inang{\rho_2}$\C{A}$\inang{\rho_2}$ $\rightarrow$
%               \C{bool}) $\rightarrow$ \C{bool}
% \end{center}
